---
title: "Study 3 - Celiktutan, Klesse & Tuk (2024)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load packages

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(emmeans)
```

## Load pretest data

```{r}
pretest <- read_excel("Data_Study3_pretest_osf.xlsx")
pretest <- pretest[pretest$Progress == 100, ]
```

## Pretest sample size

```{r}
nrow(pretest)
```

## Pretest gender

```{r}
table(pretest$gender)
prop.table(table(pretest$gender))
```

## Pretest age

```{r}
summary(pretest$age)
```

## Pretest behavior categorization

```{r}
behavior_cols <- c("B1", "B2", "B3", "B4", "B5", "B6", "B7", "B8", "B9", "B10")

for (col in behavior_cols) {
  cat(col, "\n")
  print(round(prop.table(table(pretest[[col]])) * 100, 2))
}
```

## Pretest outsourcing percentages

```{r}
outsource_pct <- data.frame(
  Behavior = 1:10,
  outsource = c(8.43, 57.83, 20.48, 68.67, 33.73, 21.69, 81.93, 90.36, 90.36, 78.31)
)

outsource_pct
```

## Pretest outsourcing bar chart

```{r}
ggplot(outsource_pct, aes(x = factor(Behavior), y = outsource)) +
  geom_col(fill = "steelblue") +
  labs(x = "Behavior", y = "Outsourcing (%)", title = "Pretest: Behaviors seen as outsourcing") +
  theme_minimal()
```

## Load main study data

```{r}
dat <- read_excel("Data_Study3_osf.xlsx")
dat <- dat[dat$Status == 0, ]
dat <- dat[dat$Progress == 100, ]
```

## Main study sample size

```{r}
nrow(dat)
```

## Main study gender

```{r}
table(dat$gender)
prop.table(table(dat$gender))
```

## Main study age

```{r}
summary(dat$age)
```

## Remove missing responses

```{r}
dat <- dat[
  dat$self != 1 |
    (!is.na(dat$B1_self) & !is.na(dat$B2_self) & !is.na(dat$B3_self) &
     !is.na(dat$B4_self) & !is.na(dat$B5_self) & !is.na(dat$B6_self) &
     !is.na(dat$B7_self) & !is.na(dat$B8_self) & !is.na(dat$B9_self) &
     !is.na(dat$B10_self)),
]

dat <- dat[
  dat$self != 0 |
    (!is.na(dat$B1_other) & !is.na(dat$B2_other) & !is.na(dat$B3_other) &
     !is.na(dat$B4_other) & !is.na(dat$B5_other) & !is.na(dat$B6_other) &
     !is.na(dat$B7_other) & !is.na(dat$B8_other) & !is.na(dat$B9_other) &
     !is.na(dat$B10_other)),
]
```

## Participants per condition

```{r}
table(dat$self)
prop.table(table(dat$self))
```

## Condition bar chart

```{r}
cond_table <- as.data.frame(table(dat$self))
names(cond_table) <- c("Condition", "Count")
cond_table$Condition <- ifelse(cond_table$Condition == 1, "Self", "Other")

ggplot(cond_table, aes(x = Condition, y = Count)) +
  geom_col(fill = "darkorange") +
  labs(title = "Participants per condition") +
  theme_minimal()
```

## Create usage likelihood variables

```{r}
dat$B1 <- ifelse(dat$self == 1, dat$B1_self, dat$B1_other)
dat$B2 <- ifelse(dat$self == 1, dat$B2_self, dat$B2_other)
dat$B3 <- ifelse(dat$self == 1, dat$B3_self, dat$B3_other)
dat$B4 <- ifelse(dat$self == 1, dat$B4_self, dat$B4_other)
dat$B5 <- ifelse(dat$self == 1, dat$B5_self, dat$B5_other)
dat$B6 <- ifelse(dat$self == 1, dat$B6_self, dat$B6_other)
dat$B7 <- ifelse(dat$self == 1, dat$B7_self, dat$B7_other)
dat$B8 <- ifelse(dat$self == 1, dat$B8_self, dat$B8_other)
dat$B9 <- ifelse(dat$self == 1, dat$B9_self, dat$B9_other)
dat$B10 <- ifelse(dat$self == 1, dat$B10_self, dat$B10_other)
```

## Reshape data to long format

```{r}
dat_long <- dat %>%
  select(ResponseId, self, B1, B2, B3, B4, B5, B6, B7, B8, B9, B10) %>%
  pivot_longer(cols = starts_with("B"), names_to = "Behavior", values_to = "Likelihood")

dat_long$Behavior <- as.numeric(gsub("B", "", dat_long$Behavior))
```

## Add outsourcing score from pretest

```{r}
dat_long$outsource <- outsource_pct$outsource[match(dat_long$Behavior, outsource_pct$Behavior)]
dat_long$outsource_centered <- dat_long$outsource - mean(dat_long$outsource)
```

## Descriptive statistics

```{r}
summary(dat_long$Likelihood)
summary(dat_long$outsource_centered)
```

## Mean likelihood by condition

```{r}
dat_long %>%
  group_by(self) %>%
  summarise(mean_likelihood = mean(Likelihood))
```

## Mean likelihood by behavior

```{r}
dat_long %>%
  group_by(Behavior) %>%
  summarise(mean_likelihood = mean(Likelihood))
```

## Mean likelihood by behavior and condition

```{r}
means_plot <- dat_long %>%
  group_by(Behavior, self) %>%
  summarise(mean_likelihood = mean(Likelihood), .groups = "drop")

means_plot$Condition <- ifelse(means_plot$self == 1, "Self", "Other")
```

## Bar chart of mean likelihood by behavior

```{r}
ggplot(means_plot, aes(x = factor(Behavior), y = mean_likelihood, fill = Condition)) +
  geom_col(position = "dodge") +
  labs(x = "Behavior", y = "Mean likelihood", title = "Usage likelihood by behavior") +
  theme_minimal()
```

## Main analysis: mixed effects model

```{r}
mixedreg <- lmer(Likelihood ~ self * outsource_centered + (1 | ResponseId), data = dat_long)
summary(mixedreg)
```

## Planned contrasts

```{r}
probe <- emmeans(mixedreg, ~ self * outsource_centered,
                 at = list(outsource_centered = c(-40, -30, -20, -10, 0, 10, 20, 30)))
contrast(probe, "revpairwise", by = "outsource_centered")
```

## Prepare data for interaction plot

```{r}
dat_long$Condition <- ifelse(dat_long$self == 1, "Self", "Other")
plot_model <- lm(Likelihood ~ Condition * outsource_centered, data = dat_long)
```

## Interaction plot data

```{r}
plotdat <- expand.grid(
  Condition = c("Self", "Other"),
  outsource_centered = seq(min(dat_long$outsource_centered), max(dat_long$outsource_centered), length.out = 100)
)

plotdat$Likelihood <- predict(plot_model, newdata = plotdat)
```

## Interaction plot

```{r}
ggplot(plotdat, aes(x = outsource_centered, y = Likelihood, color = Condition)) +
  geom_line(size = 1) +
  labs(
    x = "Outsourcing (centered)",
    y = "Usage likelihood",
    title = "Self-other difference in usage likelihood"
  ) +
  scale_color_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Boxplot of likelihood by condition

```{r}
ggplot(dat_long, aes(x = Condition, y = Likelihood, fill = Condition)) +
  geom_boxplot() +
  labs(title = "Usage likelihood: self vs other") +
  scale_fill_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "none")
```
