---
title: "Study 4 - Celiktutan, Klesse & Tuk (2024)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Load packages

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(psych)
library(lsr)
```

## Load data

```{r}
dat <- read_excel("Data_Study4_osf.xlsx")
```

## Remove test responses

```{r}
dat <- dat[dat$Q3 != "test", ]
dat <- dat[dat$Q3 != "TEST", ]
```

## Keep complete responses

```{r}
dat <- dat[dat$Status == 0, ]
dat <- dat[dat$Progress == 100, ]
dat <- dat[dat$Att_check == 1, ]
```

## Sample size

```{r}
nrow(dat)
```

## Gender

```{r}
table(dat$gender)
prop.table(table(dat$gender))
```

## Age

```{r}
summary(dat$age)
```

## Participants per condition

```{r}
table(dat$self)
prop.table(table(dat$self))
```

## Condition bar chart

```{r}
cond_table <- as.data.frame(table(dat$self))
names(cond_table) <- c("Condition", "Count")
cond_table$Condition <- ifelse(cond_table$Condition == 1, "Self", "Other")

ggplot(cond_table, aes(x = Condition, y = Count)) +
  geom_col(fill = "darkorange") +
  labs(title = "Participants per condition") +
  theme_minimal()
```

## Create inferred contribution variable

```{r}
dat$cont <- ifelse(dat$self == 1, dat$contribute_self_4, dat$cont_others_4)
```

## Create usage behavior items

```{r}
dat$inspire <- ifelse(dat$self == 1, dat$inspire_self, dat$inspire_others)
dat$copy <- ifelse(dat$self == 1, dat$copy_self, dat$copy_others)
dat$effort <- ifelse(dat$self == 1, dat$effort_self, dat$effort_others)
```

## Reliability of usage behavior items

```{r}
psych::alpha(dat[, c("inspire", "copy", "effort")])
```

## Create usage behavior composite

```{r}
dat$usagebeh <- (dat$inspire + dat$copy + dat$effort) / 3
```

## Mean inferred contribution by condition

```{r}
aggregate(cont ~ self, dat, mean)
aggregate(cont ~ self, dat, sd)
```

## Inferred contribution t-test

```{r}
t.test(cont ~ self, data = dat, var.equal = TRUE)
```

## Cohen's d for inferred contribution

```{r}
cohensD(cont ~ self, data = dat)
```

## Bar chart of inferred contribution

```{r}
cont_means <- dat %>%
  group_by(self) %>%
  summarise(mean_cont = mean(cont), .groups = "drop")

cont_means$Condition <- ifelse(cont_means$self == 1, "Self", "Other")

ggplot(cont_means, aes(x = Condition, y = mean_cont, fill = Condition)) +
  geom_col() +
  labs(y = "Inferred contribution (%)", title = "Inferred contribution by condition") +
  scale_fill_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Boxplot of inferred contribution

```{r}
dat$Condition <- ifelse(dat$self == 1, "Self", "Other")

ggplot(dat, aes(x = Condition, y = cont, fill = Condition)) +
  geom_boxplot() +
  labs(y = "Inferred contribution (%)", title = "Inferred contribution: self vs other") +
  scale_fill_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Mean usage behavior by condition

```{r}
aggregate(usagebeh ~ self, dat, mean)
aggregate(usagebeh ~ self, dat, sd)
```

## Usage behavior t-test

```{r}
t.test(usagebeh ~ self, data = dat, var.equal = TRUE)
```

## Cohen's d for usage behavior

```{r}
cohensD(usagebeh ~ self, data = dat)
```

## Bar chart of usage behavior

```{r}
usage_means <- dat %>%
  group_by(self) %>%
  summarise(mean_usage = mean(usagebeh), .groups = "drop")

usage_means$Condition <- ifelse(usage_means$self == 1, "Self", "Other")

ggplot(usage_means, aes(x = Condition, y = mean_usage, fill = Condition)) +
  geom_col() +
  labs(y = "Usage behavior (1=outsourcing, 7=inspiration)", title = "Perceived usage behavior by condition") +
  scale_fill_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Mediation: path a (self to usage behavior)

```{r}
dat$self_num <- as.numeric(dat$self)
model_a <- lm(usagebeh ~ self_num, data = dat)
summary(model_a)
```

## Mediation: path c (total effect of self on contribution)

```{r}
model_c <- lm(cont ~ self_num, data = dat)
summary(model_c)
```

## Mediation: path b and c prime (self and usage behavior on contribution)

```{r}
model_b <- lm(cont ~ self_num + usagebeh, data = dat)
summary(model_b)
```

## Indirect effect

```{r}
a <- coef(model_a)["self_num"]
b <- coef(model_b)["usagebeh"]
indirect <- a * b
indirect
```

## ChatGPT experience

```{r}
table(dat$use_chatgpt)
prop.table(table(dat$use_chatgpt))
```

## ChatGPT experience by condition

```{r}
table(dat$use_chatgpt, dat$self)
```

## Chi-square test for ChatGPT experience

```{r}
chisq.test(table(dat$use_chatgpt, dat$self))
```

## ANOVA: contribution by condition and ChatGPT experience

```{r}
dat$self_f <- factor(dat$self)
dat$use_chatgpt_f <- factor(dat$use_chatgpt)

aov_result <- aov(cont ~ self_f * use_chatgpt_f, data = dat)
summary(aov_result)
```

## Mean contribution by ChatGPT experience

```{r}
aggregate(cont ~ use_chatgpt, dat, mean)
aggregate(cont ~ use_chatgpt, dat, sd)
```

## Bar chart of contribution by ChatGPT experience

```{r}
chatgpt_means <- dat %>%
  group_by(use_chatgpt) %>%
  summarise(mean_cont = mean(cont), .groups = "drop")

chatgpt_means$Experience <- ifelse(chatgpt_means$use_chatgpt == 1, "Yes", "No")

ggplot(chatgpt_means, aes(x = Experience, y = mean_cont, fill = Experience)) +
  geom_col() +
  labs(y = "Inferred contribution (%)", title = "Contribution by ChatGPT experience") +
  scale_fill_manual(values = c("Yes" = "seagreen", "No" = "gray70")) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Create psychological ownership variable

```{r}
dat$psyown <- ifelse(dat$self == 1, dat$psyown_self, dat$psyown_others)
```

## Mean psychological ownership by condition

```{r}
aggregate(psyown ~ self, dat, mean)
aggregate(psyown ~ self, dat, sd)
```

## Psychological ownership t-test

```{r}
t.test(psyown ~ self_f, data = dat, var.equal = TRUE)
```

## Cohen's d for psychological ownership

```{r}
cohensD(psyown ~ self_f, data = dat)
```

## Bar chart of psychological ownership

```{r}
psy_means <- dat %>%
  group_by(self) %>%
  summarise(mean_psyown = mean(psyown), .groups = "drop")

psy_means$Condition <- ifelse(psy_means$self == 1, "Self", "Other")

ggplot(psy_means, aes(x = Condition, y = mean_psyown, fill = Condition)) +
  geom_col() +
  labs(y = "Psychological ownership", title = "Psychological ownership by condition") +
  scale_fill_manual(values = c("Self" = "coral", "Other" = "steelblue")) +
  theme_minimal() +
  theme(legend.position = "none")
```
